A thermodynamic view of dusty protoplanetary disks I: axisymmetric stability

### 2017-04-28 model.tex
added numerical example of unstable mode with vertically well mixed dust

\subsubsection{Vertically well-mixed dust}
Here we consider the special case where the dust-to-gas ratio is
uniform in $z$. Then the LHS of Eq. \ref{dusty_solberg2} becomes
\begin{align}
-\frac{c_s^4}{(1+\tepsilon)^4}\left(\frac{\p\ln{\rhog}}{\p
    z}\right)^2\left(\frac{\p\epsilon}{\p r}\right)^2<0.
\end{align}
Thus the disk is \emph{unstable}, since there is no effective vertical 
buoyancy to stabilize vertical motions. In fact, this situation is
identical to the pure gas, adiabtic simulations of \cite{nelson13} in
which the gas entropy is vertically uniform but has a radial
gradient. There, \citeauthor{nelson13} indeed find instability.

>>>>

\subsubsection{Vertically well-mixed dust}\label{vert_mixed_theory}
Here we consider the special case where the dust-to-gas ratio is
uniform in $z$. Then the LHS of Eq. \ref{dusty_solberg2} becomes
\begin{align}
-\frac{c_s^4}{(1+\tepsilon)^4}\left(\frac{\p\ln{\rhog}}{\p
    z}\right)^2\left(\frac{\p\epsilon}{\p r}\right)^2<0.
\end{align}
Thus the disk is \emph{unstable}, since there is no effective vertical 
buoyancy to stabilize vertical motions. This situation is
identical to the pure gas, adiabtic simulations of \cite{nelson13}
where the gas entropy is vertically uniform but has a radial
gradient. The authors indeed find instability. We give a numerical
example in the dusty context in \S\ref{vert_mixed}.

### 2017-04-28 result.tex
new section 7 on secular GI. rest of paper might need adjusting. mostly doing non-SG, but one sec and one appendix has SG
added numerical example of unstable mode with vertically well mixed dust

We present numerical solutions for  
vertically stratified disks assuming perfectly coupled dust 
($\tstop=0$). This is so that the equilibria defined in \S\ref{eqm} are exact
steady states. In this limit any growing modes must be associated with
the imposed temperature gradient (\S\ref{limits}). We are thus  
examining the effect of dust-loading on the VSI, previously studied in 
pure gas disks \citep[][\citetalias{lin15} in this section]{lin15}.

Except in  \S\ref{varHd}, we consider constant midplane
dust-to-gas ratios $\epsilon_0$ and its characteristic height
$H_\epsilon$. The latter is specified via the dust layer thickness
$\Hd(r_0)$. 
These, together with the perturbation radial wavenumber $k_x$, comprises main the
parameters of the linear problem. We choose a power-law 
midplane gas density profile, $\rho_\mathrm{g0}\propto r^p$ with $p =
-1.5$. The fiducial power-law index for the 
temperature profile is $q=-1$; and we set the gas disk aspect-ratio
$h_\mathrm{g}=0.05$. These values were also used by 
\citetalias{lin15}.

@@

>>>>

We present numerical solutions for vertically stratified,
non-self-gravitating disks assuming perfectly coupled dust
($\tstop=0$) so that the equilibria defined in \S\ref{eqm} 
are exact steady states.

Except in \S\ref{varHd}---\ref{vert_mixed}, we consider constant midplane dust-to-gas
ratios $\epsilon_0$ and characteristic thickness $H_\epsilon$. Then
the dust-to-gas ratio $\epsilon=\epsilon(z)$. In this limit any
growing modes must be associated with the imposed temperature gradient
(\S\ref{limits}). We are thus examining the effect of dust-loading on
the VSI, previously studied in pure gas disks
\citep[][\citetalias{lin15} in this section]{lin15}.

The parameters for this linear problem includes $\epsilon_0$, the dust
layer thickness $\Hd(r_0)$, and the perturbation radial wavenumber
$k_x$. We choose a power-law midplane gas density profile,
$\rho_\mathrm{g0}\propto r^p$ with $p = -1.5$. The fiducial power-law
index for the temperature profile is $q=-1$; and we set the gas disk
aspect-ratio $h_\mathrm{g}=0.05$. These values were also used by  
\citetalias{lin15}.

@@

\subsection{Pure instability with vertically well-mixed
  dust}\label{vert_mixed}

In \S\ref{vert_mixed_theory} we found that for strictly isothermal
disks, a vertically-uniform dust-to-gas ratio can be unstable
if $d\epsilon/dr\neq0$. To demonstate this numerically we consider the
same disk as above but with $q=0$ and choose $H_\mathrm{d}$ such
that $H_\epsilon=10^3\Hg$. Then vertical shear $\p_z\Omega$ is attributed
to the radially-varying dust-to-gas ratio (see Eq. \ref{vshear2}).

We show unstable modes for this disk in
Fig. \ref{vert_mixed_modes}. In order to obtain appreciable growth  
rates, we needed to consider super-Solar metalicity $Z=0.03$ and high
radial wavenumber $k_x\Hg=1800$. As expected from the discussion in
\S\ref{vert_mixed_theory}, the disk admits purely growing modes with
$\omega=0$.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/vert_mixed_modes} 
  \caption{Unstable modes in a strictly isothermal disk ($q=0$) with
    vertically uniform dust-to-gas ratio. The instability is due to
    vertical shear $\p_z\Omega\neq0$ arising from radial variations in
    the dust-to-gas ratio, $d\epsilon/dr\neq 0$. \label{vert_mixed_modes}
    }
\end{figure}
